---
title: "Autoimmune liver diseases and diabetes"
author: "Anne-Sofie H. Jensen and Marie Winther-Sorensen"
date: "2023-09-01"
output:
  html_document: default
  word_document: default
---

```{r echo=T, results='hide', warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(ukbtools)
library(stringr)
library(data.table)
library(MatchIt)
library(epitools)
library(ggplot2)
library(ggsignif)
library(ResourceSelection)
df <- fread(here("data-raw/ukb_all_datasets.csv"), data.table = FALSE)
getwd()
```

### Remove individuals who have withdrawn
```{r echo=T, results='hide', message=FALSE, warning=FALSE}
withdraw <- read_csv(here("data/Withdrawal.csv"), col_names = "eid") 
df <- anti_join(df, withdraw, by = "eid")
    ```


### Retrieving ICD diagnoses
```{r}
ukb_icd_code_meaning(icd.code = c("K732", "K754", "K738", "K743", "E10", "E11", "E14", "K70", "K760", "K703","B180","B181","B182","K746","I850","I859","I982","I983","K766","K767"), icd.version = 10)
```

### Create function to make a new dataframe only with ID's with a specific ICD10.

https://stackoverflow.com/questions/67601507/r-using-dplyr-to-find-and-filter-for-a-string-in-a-whole-data-frame
```{r echo=T, results='hide'}
find_text_filter <- function(df, tt) {
    res <- df %>%
        filter(if_any(where(is.character), ~ str_detect(.x, tt)))
    return(res)
}
```

We will now create a data frame with a column (TRUE/FALSE) for each liver disease and type of diabetes, added one by one.

Autoimmune liver diseases and cirrhosis
```{r echo=T, results='hide'}
# New dataframe only with individuals with AIH diagnosis (K73.2 + K73.8 + K75.4) and add column of AIH which is true for all these individuals

df_AIH <- df %>%
    mutate(df, AIH = TRUE) %>%
    find_text_filter("K732|K738|K754")

# Then merge df and df_AIH by ID (eid), but only add the column of AIH
df_merge <- select(df_AIH, eid, AIH) %>%
    full_join(df, df_AIH, by = "eid") # full_join leaves empty rows (i.e all that do not contain AIH remain NA)

# Add Primary biliary cholangitis K743
df_PBC <- df %>%
    mutate(df, PBC = TRUE) %>%
    find_text_filter("K743")

df_merge <- select(df_PBC, eid, PBC) %>%
    full_join(df_merge, df_PBC, by = "eid")

# Make individuals having both AIH and PBC (overlap syndrome) into a new variable (AIH.PBC)
df_merge$AIH.PBC[df_merge$AIH==T & df_merge$PBC ==T] <- TRUE
df_merge$AIH[df_merge$AIH.PBC == T] <- NA
df_merge$PBC[df_merge$AIH.PBC == T] <- NA

# Add chronic hep B
df_HBV <- df %>%
    mutate(df, HBV = TRUE) %>%
    find_text_filter("B180|B181")

df_merge <- select(df_HBV, eid, HBV) %>%
    full_join(df_merge, df_HBV, by = "eid")

# Add chronic hep C
df_HCV <- df %>%
    mutate(df, HCV = TRUE) %>%
    find_text_filter("B182") 

df_merge <- select(df_HCV, eid, HCV) %>%
    full_join(df_merge, df_HCV, by = "eid")

# Deleting those with AIH that have concurrent HBV/HCV

df_merge$AIH[df_merge$HBV == T & is.na(df_merge$HCV)] <- NA
df_merge$AIH[is.na(df_merge$HBV) & df_merge$HCV == T] <- NA
df_merge$AIH[df_merge$HBV == T & df_merge$HCV == T] <- NA

# Deleting those with PBC that have concurrent HBV/HCV

df_merge$PBC[df_merge$HBV == T & is.na(df_merge$HCV)] <- NA
df_merge$PBC[is.na(df_merge$HBV) & df_merge$HCV == T] <- NA

# Cirrhosis
df_cir <- df %>%
    mutate(df, cir = TRUE) %>%
    find_text_filter("K746|I850|I859|I982|I983|K729|K703|K766|K767")

df_merge <- select(df_cir, eid, cir) %>%
    full_join(df_merge, df_cir, by = "eid")

# Stringent cirrhosis definition
df_cir2 <- df %>%
    mutate(df, cir2 = TRUE) %>%
    find_text_filter("K746|K703")

df_merge <- select(df_cir2, eid, cir2) %>%
    full_join(df_merge, df_cir2, by = "eid")

```


Type 2 diabetes: E11+E14
Type 1 diabetes: E10
Exclusion: if only gestational diabetes => no diabetes

```{r echo=TRUE, results = 'hide'}
# T2D
df_T2D <- df %>%
    mutate(df, T2D = TRUE) %>%
    find_text_filter("E11|E14")

df_merge <- select(df_T2D, eid, T2D) %>%
    full_join(df_merge, df_T2D, by = "eid")

# T1D
df_T1D <- df %>%
    mutate(df, T1D = TRUE) %>%
    find_text_filter("E10")

df_merge <- select(df_T1D, eid, T1D) %>%
    full_join(df_merge, df_T1D, by = "eid")

# Remove diabetes diagnosis if gestational diabetes only
df_merge$gdmonly[df_merge$gestational_diabetes_only_f4041_0_0 == 1 | df_merge$gestational_diabetes_only_f4041_1_0 == 1 | df_merge$gestational_diabetes_only_f4041_2_0 == 1 | df_merge$gestational_diabetes_only_f4041_3_0 == 1] <- T
df_merge$T1D[df_merge$gdmonly == T & df_merge$T2D == T] <- NA
df_merge$T2D[df_merge$gdmonly == T & is.na(df_merge$T1D)] <- NA
df_merge$T1D[df_merge$gdmonly == T & is.na(df_merge$T2D)] <- NA

```

### Diabetes medication codes
#### Metformin
1140884600	metformin
1140874686  glucophage 500 mg tablet 
1141189090	rosiglitazone 1mg / metformin 500mg tablet

#### Sulfonylureas
1140874718  glibenclamide 
1140874744	gliclazide
1140874746  diamicron 80 mg tablet 
1141152590  glimepiride
1141156984  amaryl 1 mg tablet
1140874646	glipizide
1141157284	glipizide product
1140874652  minodiab 2.5 mg tablet
1140874674  tolbutamide
1140874728  euglucon 2.5 mg

#### Others
1140868902	acarbose
1140868908  glucobay 50 mg tablet
1140857508  glucotard 5g/sachet mini-tablet
1140882964  oral hypoglycemic

#### Metglitinides
1141173882	nateglinide
1141173786  starlix 60 mg tablet
1141168660	repaglinide

#### Glitazones
1141171646	pioglitazone
1141171652  actos 15 mg tablet
1141153254  troglitazone
1141177600	rosiglitazone
1141177606  avandia 4mg tablet

#### Insulin
1140883066	insulin product

### Filter for oral antihyperglycemic drugs and insulin
```{r echo=TRUE, results='hide'}
# Since the columns with medication codes (treatmentmedication_code_f20003_0_i) are integers, we create a new function taking integers as inputs:

find_drug_function <- function(df, tt) {
    res <- df %>%
        filter(if_any(where(is.integer), ~ str_detect(.x, tt)))
    return(res)}

# Oral T2D treatment
df_oral_T2D_drug <- df %>%
    mutate(df, oral_T2D_drug = TRUE) %>%
    find_drug_function("1140874718|1140874744|1140874746|1141152590|1141156984|1140874646|1141157284|1140874652|1140874674|1140874728|1140868902|1140868908|1140857508|1141173882|1141173786|1141168660|1141171646|1141171652|1141153254|1141177600|1141177606|1140882964")
    
df_merge <- select(df_oral_T2D_drug, eid, oral_T2D_drug) %>%
    full_join(df_merge, df_oral_T2D_drug, by = "eid")

# Metformin treatment
df_metformin <- df %>%
    mutate(df, metformin = TRUE) %>%
    find_drug_function("1140884600|1140874686|1141189090") 

df_merge <- select(df_metformin, eid, metformin) %>%
    full_join(df_merge, df_metformin, by = "eid")

# Insulin treatment
df_insulin <- df %>%
    mutate(df, insulin_treatment = TRUE) %>%
    find_drug_function("1140883066")

df_merge <- select(df_insulin, eid, insulin_treatment) %>%
    full_join(df_merge, df_insulin, by = "eid")


```

### Calculate sum of average weekly intake of alcohol

```{r echo=TRUE, results='hide'}
# Set prefer not to answer and do not know (-1 and -3) to NA
df_merge$average_weekly_beer_plus_cider_intake_f1588_0_0[df_merge$average_weekly_beer_plus_cider_intake_f1588_0_0<0] <- NA
df_merge$average_weekly_champagne_plus_white_wine_intake_f1578_0_0[df_merge$average_weekly_champagne_plus_white_wine_intake_f1578_0_0<0] <- NA
df_merge$average_weekly_fortified_wine_intake_f1608_0_0[df_merge$average_weekly_fortified_wine_intake_f1608_0_0<0] <- NA
df_merge$average_weekly_intake_of_other_alcoholic_drinks_f5364_0_0[df_merge$average_weekly_intake_of_other_alcoholic_drinks_f5364_0_0<0] <- NA
df_merge$average_weekly_red_wine_intake_f1568_0_0[df_merge$average_weekly_red_wine_intake_f1568_0_0<0] <- NA
df_merge$average_weekly_spirits_intake_f1598_0_0[df_merge$average_weekly_spirits_intake_f1598_0_0<0] <- NA

# units of alcohol in beverages: https://www.nhs.uk/live-well/alcohol-advice/calculating-alcohol-units/

# weekly units of alcohol - redwine (f1568), 1 glass = 2.1 units
df_merge$average_weekly_red_wine_intake_f1568_0_0 <- df_merge$average_weekly_red_wine_intake_f1568_0_0 * 2.1
# weekly units of alcohol - champagne etc (f1578), 1 glass = 2.1 units
df_merge$average_weekly_champagne_plus_white_wine_intake_f1578_0_0 <- df_merge$average_weekly_champagne_plus_white_wine_intake_f1578_0_0 * 2.1
# weekly units of alcohol - beer and cider (f1588), 1 pint = 2 units
df_merge$average_weekly_beer_plus_cider_intake_f1588_0_0 <- df_merge$average_weekly_beer_plus_cider_intake_f1588_0_0 * 2
# weekly units of alcohol - spirits (f1598), 1 measure = 1 unit (no calculation needed)
# weekly units of alcohol - fortified wine (f1608), 1 measure = 1 unit (no calculation needed)
# weekly units of other alcoholic drinks incl alcopops, 1 glass = 1.5 units
df_merge$average_weekly_intake_of_other_alcoholic_drinks_f5364_0_0 <- df_merge$average_weekly_intake_of_other_alcoholic_drinks_f5364_0_0 * 1.5

#Calculate the sum of weekly units alcoholic intake
df_merge <- df_merge %>%
    rowwise() %>%
    mutate(weekly_alcohol = sum(average_weekly_beer_plus_cider_intake_f1588_0_0,
    average_weekly_champagne_plus_white_wine_intake_f1578_0_0,
    average_weekly_fortified_wine_intake_f1608_0_0,
    average_weekly_intake_of_other_alcoholic_drinks_f5364_0_0,
    average_weekly_red_wine_intake_f1568_0_0,
    average_weekly_spirits_intake_f1598_0_0,
    na.rm=T))

```

### Defining NAFLD
```{r echo =TRUE, results='hide'}
# NAFLD
df_NAFLD <- df %>%
    mutate(df, NAFLD = TRUE) %>%
    find_text_filter("K760")

df_merge <- select(df_NAFLD, eid, NAFLD) %>%
    full_join(df_merge, df_NAFLD, by = "eid")

# Number of PDFF measures in dataset with 40534 variable
table(is.na(df_merge$proton_density_fat_fraction_pdff_f40061_2_0))

# Add pdff > 5.5 for women and men based on alcohol intake (NAFLD when drinking less than 140 g alcohol per week for women and less than 210 g alcohol per week for men (EASL guideline 2016). 1 unit = 8 g alcohol => NAFLD when less than < 17.5 units alcohol/week for women and 26.25 units alcohol/week for men
df_merge$NAFLD[df_merge$proton_density_fat_fraction_pdff_f40061_2_0 > 5.5 &
                 df_merge$sex_f31_0_0 == "Female" &
                 df_merge$weekly_alcohol < 17.5] <- T
df_merge$NAFLD[df_merge$proton_density_fat_fraction_pdff_f40061_2_0 > 5.5 &
                 df_merge$sex_f31_0_0 == "Male" &
                 df_merge$weekly_alcohol<26.25] <- T

```


### Defining alcoholic liver diseases
```{r echo=TRUE, results='hide'}
# Add ALD K70
df_ALD <- df %>%
    mutate(df, ALD = TRUE) %>%
    find_text_filter("K70")

df_merge <- select(df_ALD, eid, ALD) %>%
    full_join(df_merge, ALD, by = "eid")

# Add pdff > 5.5 for women and men based on alcohol intake
df_merge$ALD[df_merge$proton_density_fat_fraction_pdff_f40061_2_0 > 5.5 &
                 df_merge$sex_f31_0_0 == "Female" &
                 df_merge$weekly_alcohol >= 17.5] <- T
df_merge$ALD[df_merge$proton_density_fat_fraction_pdff_f40061_2_0 > 5.5 &
                 df_merge$sex_f31_0_0=="Male" &
                 df_merge$weekly_alcohol >= 26.25] <- T

```


### Tidy up dataframe
```{r echo=TRUE, results='hide'}
# Make NA -> FALSE
df_final <- df_merge %>%
    replace_na(list(
        T1D = FALSE,
        T2D = FALSE,
        AIH = FALSE,
        PBC = FALSE,
        AIH.PBC = FALSE,
        NAFLD = FALSE,
        ALD = FALSE,
        hh = FALSE,
        insulin_treatment = FALSE,
        oral_T2D_drug = FALSE,
        metformin = FALSE,
        cir = FALSE,
        cir2 = FALSE)) # Make NA's FALSE

# Clean up dataframe
df_final <- df_final %>%
    select(ID = eid,
           Age = age_when_attended_assessment_centre_f21003_0_0,
           Sex = sex_f31_0_0,
           hba1c = glycated_haemoglobin_hba1c_f30750_0_0,
           BMI = body_mass_index_bmi_f21001_0_0,
           AIH,
           PBC,
           AIH.PBC,
           NAFLD,
           cir,
           cir2,
           ALD,
           T1D,
           T2D,
           insulin_treatment,
           oral_T2D_drug,
           metformin,
           age_diabetes_diagnosed_f2976_0_0,
           age_diabetes_diagnosed_f2976_1_0,
           age_diabetes_diagnosed_f2976_2_0,
           age_diabetes_diagnosed_f2976_3_0,
           started_insulin_within_one_year_diagnosis_of_diabetes_f2986_0_0,
           started_insulin_within_one_year_diagnosis_of_diabetes_f2986_1_0,
           started_insulin_within_one_year_diagnosis_of_diabetes_f2986_2_0,
           started_insulin_within_one_year_diagnosis_of_diabetes_f2986_3_0,
           ethnicity = ethnic_background_f21000_0_0)

```

### Defining diabetes

```{r echo = TRUE, results='hide'}

# MAke those individuals wihtout diabetes diagnosis as T2D if hba1c above 48:
df_final$T2D[df_final$hba1c>48 & (df_final$T2D == F & df_final$T1D == F)] <-  T

# Make those individuals without diabetes diagnosis as T2D if oral T2D medication (excluding metformin):
df_final$T2D[df_final$oral_T2D_drug == T & (df_final$T2D == F & df_final$T1D == F)] <- T

# Make those T1D individuals on oral T2D drug medication T2D:
df_final$T1D[(df_final$T2D == T & df_final$T1D == T) & (df_final$oral_T2D_drug == T | df_final$metformin == T)] <- F
df_final$T2D[df_final$T1D == T & (df_final$oral_T2D_drug == T | df_final$metformin == T)] <- T
df_final$T1D[df_final$oral_T2D_drug == T | df_final$metformin == T] <- F

# Make those diagnosed <20 y age as T1D
df_final$T2D[(df_final$age_diabetes_diagnosed_f2976_0_0 < 20 | df_final$age_diabetes_diagnosed_f2976_1_0 < 20 | df_final$age_diabetes_diagnosed_f2976_2_0 < 20 | df_final$age_diabetes_diagnosed_f2976_3_0 < 20 ) & df_final$T1D == T & df_final$T2D == T] <- F

# Make those with double diagnosis as T1D if insulin treatment within 1 year of diagnosis and actual insulin treatment:
df_final$ins1y[df_final$started_insulin_within_one_year_diagnosis_of_diabetes_f2986_0_0 == 1 | df_final$started_insulin_within_one_year_diagnosis_of_diabetes_f2986_1_0 == 1 | df_final$started_insulin_within_one_year_diagnosis_of_diabetes_f2986_2_0 == 1 | df_final$started_insulin_within_one_year_diagnosis_of_diabetes_f2986_3_0 == 1] <- T 
df_final$ins1y[is.na(df_final$ins1y)] <- F
df_final$T2D[df_final$T1D == T & df_final$T2D == T & (df_final$ins1y & df_final$insulin_treatment)] <- F

# Making those with double diagnosis T1D and T2D as T2D
df_final$T1D[df_final$T1D == T & df_final$T2D == T] <- F

# Prevalence estimate (n(diagnosis)/total count)
count(df_final)
table(df_final$AIH)
table(df_final$PBC)
table(df_final$AIH.PBC)
table(df_final$NAFLD)
table(df_final$ALD)
table(df_final$T1D)
table(df_final$cir)
table(df_final$NAFLD)

df_final %>%
      count(T1D, T2D)

```

### Ethnicity

```{r echo=TRUE, results='hide'}
# All ethnic groups
df_final$ethnic_all <- df_final$ethnicity
df_final$ethnic_all[df_final$ethnic_all=="Caribbean"] <- 4001
df_final$ethnic_all[df_final$ethnic_all=="Indian"] <- 3001
df_final$ethnic_all[df_final$ethnic_all=="White"] <- 1
df_final$ethnic_all[df_final$ethnic_all=="White and Black Caribbean"] <- 2001
df_final$ethnic_all[df_final$ethnic_all=="British"] <- 1001
df_final$ethnic_all[df_final$ethnic_all=="Pakistani"] <- 3002
df_final$ethnic_all[df_final$ethnic_all=="Mixed"] <- 2
df_final$ethnic_all[df_final$ethnic_all=="African"] <- 4002
df_final$ethnic_all[df_final$ethnic_all=="Irish"] <- 1002
df_final$ethnic_all[df_final$ethnic_all=="White and Black African"] <- 2002
df_final$ethnic_all[df_final$ethnic_all=="Bangladeshi"] <- 3003
df_final$ethnic_all[df_final$ethnic_all=="Asian or Asian British"] <- 3
df_final$ethnic_all[df_final$ethnic_all=="Any other Black background"] <- 4003
df_final$ethnic_all[df_final$ethnic_all=="Any other white background"] <- 1003
df_final$ethnic_all[df_final$ethnic_all=="White and Asian"] <- 2003
df_final$ethnic_all[df_final$ethnic_all=="Any other Asian background"] <- 3004
df_final$ethnic_all[df_final$ethnic_all=="Black or Black British"] <- 4
df_final$ethnic_all[df_final$ethnic_all=="Any other mixed background"] <- 2004
df_final$ethnic_all[df_final$ethnic_all=="Chinese"] <- 5
df_final$ethnic_all[df_final$ethnic_all=="Other ethnic group"] <- 6
df_final$ethnic_all[df_final$ethnic_all=="Do not know"] <- -1
df_final$ethnic_all[df_final$ethnic_all=="Prefer not to answer"] <- -3

#Main ethnic groups (NB this makes some assumptions i.e. British is white, and does not include mixed, Chinese or other Asian)

df_final$ethnic <- df_final$ethnic_all
df_final$ethnic[df_final$ethnic_all==1001] <- 1
df_final$ethnic[df_final$ethnic_all==1002] <- 1
df_final$ethnic[df_final$ethnic_all==1003] <- 1
df_final$ethnic[df_final$ethnic_all==3] <- 2
df_final$ethnic[df_final$ethnic_all==3001] <- 2
df_final$ethnic[df_final$ethnic_all==3002] <- 2
df_final$ethnic[df_final$ethnic_all==3003] <- 2
df_final$ethnic[df_final$ethnic_all==4] <- 3
df_final$ethnic[df_final$ethnic_all==4001] <- 3
df_final$ethnic[df_final$ethnic_all==4002] <- 3
df_final$ethnic[df_final$ethnic_all==4003] <- 3
df_final$ethnic[df_final$ethnic_all==2] <- 4
df_final$ethnic[df_final$ethnic_all==2001] <- 4
df_final$ethnic[df_final$ethnic_all==2002] <- 4
df_final$ethnic[df_final$ethnic_all==2003] <- 4
df_final$ethnic[df_final$ethnic_all==2004] <- 4
df_final$ethnic[df_final$ethnic_all==3004] <- 4
df_final$ethnic[df_final$ethnic_all==5] <- 4
df_final$ethnic[df_final$ethnic_all==6] <- 4
df_final$ethnic[df_final$ethnic_all==-1] <- NA
df_final$ethnic[df_final$ethnic_all==-3] <- NA

count(df_final,ethnic_all) # empty row, should be NA
df_final$ethnic_all # empty row = ""

df_final$ethnic[df_final$ethnic_all==""] <- NA

```
### Exclude participants with no registration of age, sex, BMI, ethnicity

```{r echo=TRUE, results = 'hide'}
# As matchit function does not allow NAs in covariats (age, sex, BMI, ethnicity), these have to be deleted. Create new dataframe df_merge_match

df_merge_match <- df_final
df_merge_match <- df_merge_match[!is.na(df_merge_match[,"Age"]),]
df_merge_match <- df_merge_match[!is.na(df_merge_match[,"Sex"]),]
df_merge_match <- df_merge_match[!is.na(df_merge_match[,"BMI"]),]
df_merge_match <- df_merge_match[!is.na(df_merge_match[,"ethnic"]),]

```

### Count and baseline characteristics

```{r echo=TRUE, results='hide'}
# Count of diagnoses after age, sex, BMI, ethnicity exclusion
count(df_final)
count(df_merge_match)
table(df_merge_match$AIH)
table(df_merge_match$PBC)
table(df_merge_match$AIH.PBC)
table(df_merge_match$NAFLD)
table(df_merge_match$ALD)
table(df_merge_match$T1D)

# Baseline characteristics of AIH, PBC and AIH-PBC overlap syndrome, NAFLD, ALD

# AIH
summary(df_merge_match$Age[df_merge_match$AIH == T])
sd(df_merge_match$Age[df_merge_match$AIH == T])
table(df_merge_match$Sex[df_merge_match$AIH == T])
summary(df_merge_match$BMI[df_merge_match$AIH == T])
sd(df_merge_match$BMI[df_merge_match$AIH == T],na.rm = T)
table(df_merge_match$AIH,df_merge_match$ethnic)

# PBC
summary(df_merge_match$Age[df_merge_match$PBC == T])
sd(df_merge_match$Age[df_merge_match$PBC == T])
table(df_merge_match$Sex[df_merge_match$PBC == T])
summary(df_merge_match$BMI[df_merge_match$PBC == T])
sd(df_merge_match$BMI[df_merge_match$PBC == T],na.rm = T)
table(df_merge_match$PBC,df_merge_match$ethnic)

# AIH-PBC overlap syndrome
summary(df_merge_match$Age[df_merge_match$AIH.PBC == T])
sd(df_merge_match$Age[df_merge_match$AIH.PBC == T])
table(df_merge_match$Sex[df_merge_match$AIH.PBC == T])
summary(df_merge_match$BMI[df_merge_match$AIH.PBC == T])
sd(df_merge_match$BMI[df_merge_match$AIH.PBC == T],na.rm = T)
table(df_merge_match$AIH.PBC,df_merge_match$ethnic)

# NAFLD
summary(df_merge_match$Age[df_merge_match$NAFLD == T])
sd(df_merge_match$Age[df_merge_match$NAFLD == T])
table(df_merge_match$Sex[df_merge_match$NAFLD == T])
summary(df_merge_match$BMI[df_merge_match$NAFLD == T])
sd(df_merge_match$BMI[df_merge_match$NAFLD == T],na.rm = T)
table(df_merge_match$NAFLD,df_merge_match$ethnic)

# ALD
summary(df_merge_match$Age[df_merge_match$ALD == T])
sd(df_merge_match$Age[df_merge_match$ALD == T])
table(df_merge_match$Sex[df_merge_match$ALD == T])
summary(df_merge_match$BMI[df_merge_match$ALD == T])
sd(df_merge_match$BMI[df_merge_match$ALD == T],na.rm = T)
table(df_merge_match$ALD,df_merge_match$ethnic)

```


## Matched controls (age, sex, BMI, ethnicity) in AIH:

Looking at the distribution of covariats before matching:

```{r echo = TRUE, results='hide',fig.keep='none'}
df.m.0.aih <- matchit(AIH ~ Age + Sex + BMI + ethnic, data = df_merge_match, method = NULL)
summary(df.m.0.aih)
plot(summary(df.m.0.aih))
```

Now matching (5 matched controls per AIH patient) - nearest neighboor propensity score matching method:

```{r echo = TRUE, results='hide',fig.keep='none'}
df.m.aih <- matchit(AIH ~ Age + Sex + BMI + ethnic, data=df_merge_match, ratio = 5)
summary(df.m.aih)
plot(summary(df.m.aih))
df.m.aih.data <- match.data(df.m.aih) # creating new dataset
```

T2D risk in AIH:

```{r echo=TRUE, results='hide'}
table(df.m.aih.data$AIH, df.m.aih.data$T2D)

glm.aih.t2d <- glm(data = df.m.aih.data, T2D ~ AIH, family = "binomial")
summary(glm.aih.t2d)
exp(coef(glm.aih.t2d)) # OR
exp(confint.default(glm.aih.t2d)) # CI

```

Is risk caused by cirrhosis/NAFLD:

```{r echo=TRUE, results='hide'}
# matching on age, sex, BMI, ethnicity, NAFLD

df.m.aih.nafld <- matchit(AIH ~ Age + Sex + BMI + ethnic + NAFLD, data=df_merge_match, ratio = 5)
df.m.aih.nafld.data <- match.data(df.m.aih.nafld)
table(df.m.aih.nafld.data$AIH)

glm.aih.t2d.nafld <- glm(data = df.m.aih.nafld.data, T2D ~ AIH, family = "binomial")
summary(glm.aih.t2d.nafld)
exp(coef(glm.aih.t2d.nafld))
exp(confint.default(glm.aih.t2d.nafld))

# matching on age, sex, BMI, ethnicity, cirrhosis (broad cirrhosis definition)

df.m.aih.cir <- matchit(AIH ~ Age + Sex + BMI + ethnic + cir, data=df_merge_match, ratio = 5)
df.m.aih.cir.data <- match.data(df.m.aih.cir)
table(df.m.aih.cir.data$AIH)

glm.aih.t2d.cir <- glm(data = df.m.aih.cir.data, T2D ~AIH, family = "binomial")
summary(glm.aih.t2d.cir)
exp(coef(glm.aih.t2d.cir))
exp(confint.default(glm.aih.t2d.cir))

# matching on age, sex, BMI, ethnicity, cirrhosis (stringent cirrhosis definition)

df.m.aih.cir2 <- matchit(AIH ~ Age + Sex + BMI + ethnic + cir2, data=df_merge_match, ratio = 5)
df.m.aih.cir2.data <- match.data(df.m.aih.cir2)

glm.aih.t2d.cir2 <- glm(data = df.m.aih.cir2.data, T2D ~AIH, family = "binomial")
summary(glm.aih.t2d.cir2)
exp(coef(glm.aih.t2d.cir2))
exp(confint.default(glm.aih.t2d.cir2))

```
T1D risk in AIH:

```{r echo=TRUE, results='hide',warning=FALSE}
table(df.m.aih.data$AIH, df.m.aih.data$T1D)
chisq.test(table(df.m.aih.data$AIH,df.m.aih.data$T1D)) $ expected # As expected count below 5 => fisher's test
fisher.test(table(df.m.aih.data$AIH, df.m.aih.data$T1D))
```

## Matched controls (age, sex, BMI, ethnicity) in PBC:

Looking at the distribution of covariats before matching:

```{r echo=TRUE, results='hide',fig.keep='none'}
df.m.0.pbc <- matchit(PBC ~ Age + Sex + BMI + ethnic, data = df_merge_match, method = NULL)
summary(df.m.0.pbc) 
plot(summary(df.m.0.pbc))
```

Now matching (5 matched controls per PBC patient):

```{r echo = TRUE, results='hide',fig.keep='none'}
df.m.pbc <- matchit(PBC ~ Age + Sex + BMI + ethnic, data=df_merge_match, ratio = 5)
summary(df.m.pbc)
plot(summary(df.m.aih))
df.m.pbc.data <- match.data(df.m.pbc)
```

T2D risk in PBC:

```{r echo = TRUE, results='hide'}
glm.pbc.t2d <- glm(data = df.m.pbc.data, T2D ~ PBC, family = "binomial")
summary(glm.pbc.t2d)
exp(coef(glm.pbc.t2d)) # OR
exp(confint.default(glm.pbc.t2d)) # CI

```

Is the increased risk caused by cirrhosis / NAFLD :

```{r echo=TRUE, results='hide'}
# matching on age, sex, BMI, ethnicity, NAFLD

df.m.pbc.nafld <- matchit(PBC ~ Age + Sex + BMI + ethnic + NAFLD, data=df_merge_match, ratio = 5)
df.m.pbc.nafld.data <- match.data(df.m.pbc.nafld)

glm.pbc.t2d.nafld <- glm(data = df.m.pbc.nafld.data, T2D ~PBC, family = "binomial")
summary(glm.pbc.t2d.nafld)
exp(coef(glm.pbc.t2d.nafld)) 
exp(confint.default(glm.pbc.t2d.nafld))

# matching on age, sex, BMI, ethnicity, cirrhosis (broad cirrhosis definition)

df.m.pbc.cir <- matchit(PBC ~ Age + Sex + BMI + ethnic + cir, data=df_merge_match, ratio = 5)
df.m.pbc.cir.data <- match.data(df.m.pbc.cir)

glm.pbc.t2d.cir <- glm(data = df.m.pbc.cir.data, T2D ~PBC, family = "binomial")
summary(glm.pbc.t2d.cir)
exp(coef(glm.pbc.t2d.cir))
exp(confint.default(glm.pbc.t2d.cir))

# matching on age, sex, BMI, ethnicity, cirrhosis (stringent cirrhosis definition)

df.m.pbc.cir2 <- matchit(PBC ~ Age + Sex + BMI + ethnic + cir2, data=df_merge_match, ratio = 5)
df.m.pbc.cir2.data <- match.data(df.m.pbc.cir2)

glm.pbc.t2d.cir2 <- glm(data = df.m.pbc.cir2.data, T2D ~PBC, family = "binomial")
summary(glm.pbc.t2d.cir2)
exp(coef(glm.pbc.t2d.cir2))
exp(confint.default(glm.pbc.t2d.cir2))

```

T1D risk in PBC:

```{r echo=TRUE, results='hide',warning=FALSE}
table(df.m.pbc.data$PBC, df.m.pbc.data$T1D)
chisq.test(table(df.m.pbc.data$PBC,df.m.pbc.data$T1D)) $ expected # As expected count below 5 => fisher's test
fisher.test(table(df.m.pbc.data$PBC, df.m.pbc.data$T1D))
```

## Matched controls (age, sex, BMI, ethnicity) in AIH-PBC overlap:

Looking at the distribution of covariats before matching:

```{r echo=TRUE, results='hide',fig.keep='none'}
df.m.0.aihpbc <- matchit(AIH.PBC ~ Age + Sex + BMI + ethnic, data = df_merge_match, method = NULL)
summary(df.m.0.aihpbc) 
plot(summary(df.m.0.aihpbc))
```

Now matching (5 matched controls per AIH-PBC patient):

```{r echo=TRUE, results='hide',fig.keep='none'}
df.m.aihpbc <- matchit(AIH.PBC ~ Age + Sex + BMI + ethnic, data=df_merge_match, ratio = 5)
summary(df.m.aihpbc)
plot(summary(df.m.aihpbc))
df.m.aihpbc.data <- match.data(df.m.aihpbc)
```

T2D risk in AIH-PBC:

```{r echo = TRUE, results='hide'}

glm.aihpbc.t2d <- glm(data = df.m.aihpbc.data, T2D ~ AIH.PBC, family = "binomial")
summary(glm.aihpbc.t2d)
exp(coef(glm.aihpbc.t2d)) # OR
exp(confint.default(glm.aihpbc.t2d)) # CI

```


T1D risk in AIH-PBC:

```{r echo=TRUE, results='hide',warning=FALSE}
table(df.m.aihpbc.data$AIH.PBC, df.m.aihpbc.data$T1D)
chisq.test(table(df.m.aihpbc.data$AIH.PBC,df.m.aihpbc.data$T1D)) $ expected # As expected count below 5 => fisher's test
fisher.test(table(df.m.aihpbc.data$AIH.PBC, df.m.aihpbc.data$T1D))

```

## Matched controls (age, sex, BMI, ethnicity) in ALD:

Looking at the distribution of covariats before matching:

```{r echo=TRUE, results='hide',fig.keep='none'}
df.m.0.ald <- matchit(ALD ~ Age + Sex + BMI + ethnic, data = df_merge_match, method = NULL)
summary(df.m.0.ald)
plot(summary(df.m.0.ald))
```

Now matching (5 matched controls per ALD): 

```{r echo=TRUE, results='hide',fig.keep='none'}
df.m.ald <- matchit(ALD ~ Age + Sex + BMI + ethnic, data=df_merge_match, ratio = 5)
summary(df.m.ald)
plot(summary(df.m.ald))
df.m.ald.data <- match.data(df.m.ald)
```

T2D risk in ALD:
 
```{r echo=TRUE, results='hide'}
glm.ald.t2d <- glm(data = df.m.ald.data, T2D ~ ALD, family = "binomial")
summary(glm.ald.t2d)
exp(coef(glm.ald.t2d)) # OR
exp(confint.default(glm.ald.t2d)) # CI
```

## Matched controls (age, sex, BMI, ethnicity) in NAFLD:

Looking at the distribution of covariats before matching:

```{r echo=TRUE, results='hide',fig.keep='none'}
df.m.0.nafld <- matchit(NAFLD ~ Age + Sex + BMI + ethnic, data = df_merge_match, method = NULL)
summary(df.m.0.nafld)
plot(summary(df.m.0.nafld))
```

Now matching (5 matched controls per NAFLD): 

```{r echo=TRUE, results='hide',fig.keep='none'}
df.m.nafld <- matchit(NAFLD ~ Age + Sex + BMI + ethnic, data=df_merge_match, ratio = 5)
summary(df.m.nafld) 
plot(summary(df.m.nafld))
df.m.nafld.data <- match.data(df.m.nafld)
```

T2D risk in NAFLD:
 
```{r echo=TRUE, results='hide'}
glm.nafld.t2d <- glm(data = df.m.nafld.data, T2D ~ NAFLD, family = "binomial")
summary(glm.nafld.t2d)
exp(coef(glm.nafld.t2d)) # OR
exp(confint.default(glm.nafld.t2d)) # CI

```

## Barplot T2D

```{r}
# Create dataframe

diagnosis <- rep(c("AIH","PBC","AIH-PBC","NAFLD","ALD"),each=2)
Proportion <- rep(c("Cases","Age-,sex-, BMI-, and ethnicity-matched controls"),times=5)
value <- c(13.7,8.6,15.7,9.1,14.6,8.8,22.2,12.2,14.7,10.5)

T2D.df <- data.frame(diagnosis,Proportion,value)

# To correct X-axis order (else alphabetical order) =>
T2D.df$diagnosis <- factor(T2D.df$diagnosis,levels=unique(T2D.df$diagnosis)) 
T2D.df$Proportion <- factor(T2D.df$Proportion,levels=unique(T2D.df$Proportion))

ggplot(T2D.df, aes(x = diagnosis, y = value)) +  
  geom_bar(stat = 'identity', position = 'dodge',aes(fill = Proportion)) + 
  scale_fill_manual(values = c("#5B9BD5", "#44546A")) + 
  ylab("Proportion type 2 diabetes (%)") + 
  xlab("Diagnosis") + 
  labs(fill = "") + 
  geom_signif(stat = "identity", data = data.frame(x = c(0.7, 1.7, 2.7,3.7,4.7), 
                                                   xend = c(1.3, 2.3, 3.3,4.3,5.3), 
                                                   y = c(14.7, 16.7, 15.6,23.2,15.7), 
                                                   annotation = c("")), 
              aes(x = x, xend = xend, y=y, yend = y, annotation = annotation)) + 
  annotate("text", x = 1:5, y = c(15.7, 17.7, 16.6,24.2,16.7), label = c("**","***","NS","***","***")) +
  geom_text(aes (label = paste(format(value,nsmall = 1))), nudge_x = c(-0.225,0.225), nudge_y = -1, col = "white")

# * = p-value < 0.05, ** = p-value < 0.01, *** = p-value < 0.001

```

